Deep mutational scanning of hemagglutinin helps distinguish the evolutionary fate of human H3N2 influenza virus lineages

### 2018-01-03 paper/manuscript.tex
Start moving manuscript over to _PNAS_ LaTex template

% Use only LaTeX2e, calling the article.cls class and 12-point type.

\documentclass[11pt]{article}
\usepackage[round,semicolon]{natbib}
\usepackage[margin=1.25in]{geometry}
\usepackage{kpfonts}

\usepackage{seqsplit}
\usepackage{placeins}

\usepackage{newfloat}
\usepackage[labelfont=bf]{caption}
\usepackage{nameref}
\usepackage{rotating}
\usepackage{color}
\usepackage{float}

\setcounter{topnumber}{8}
\setcounter{bottomnumber}{8}
\setcounter{totalnumber}{8}
\renewcommand{\topfraction}{1}
\renewcommand{\bottomfraction}{1}
\renewcommand{\textfraction}{0}
\renewcommand{\floatpagefraction}{1}

\usepackage[font=small,labelfont=bf]{caption}

\usepackage{newfloat}
\DeclareFloatingEnvironment[name={Figure}]{suppfigure}
\renewcommand{\thesuppfigure}{S\arabic{suppfigure}}
\DeclareFloatingEnvironment[name={Table}]{supptable}
\renewcommand{\thesupptable}{S\arabic{supptable}}
\DeclareFloatingEnvironment[name={File}]{suppfile}
\renewcommand{\thesuppfile}{S\arabic{suppfile}}

\definecolor{darkblue}{rgb}{0, 0.0, 0.6}

\usepackage{hyperref}
\hypersetup{colorlinks,citecolor=blue,linkcolor=blue,urlcolor=blue}

\usepackage{seqsplit}

\usepackage{array}
\newcolumntype{R}[1]{>{\raggedright\arraybackslash}p{#1}}
\newcolumntype{C}[1]{>{\centering\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}

\newcommand{\comment}[1]{{\color{red}[\textsl{#1}]}}

\usepackage{setspace}

\renewcommand{\topfraction}{1}
\renewcommand{\bottomfraction}{1}
\renewcommand{\textfraction}{0}
\renewcommand{\floatpagefraction}{1}

@@

\author{
Juhye M. Lee$^{1,4,5,\dagger}$, \and 
John Huddleston$^{2,6,\dagger}$, \and 
Michael B. Doud$^{1,4,5}$, \and 
Kathryn A. Hooper$^{1,6}$, \and
Trevor Bedford$^{2,3*}$ and \and 
Jesse D. Bloom$^{1,3,4*}$
\\
\scriptsize{$^1$Basic Sciences Division, $^2$Vaccine and Infectious Diseases Division, and $^3$Computational Biology Program,} \\
\scriptsize{Fred Hutchinson Cancer Research Center, Seattle, WA, USA} \\
\scriptsize{$^4$Department of Genome Sciences, $^5$Medical Scientist Training Program, and $^6$Molecular and Cellular Biology Program,} \\
\scriptsize{University of Washington, Seattle, WA, USA} \\
\scriptsize{$^{\dagger}$These authors contributed equally} \\
\scriptsize{$^*$Correspondence: \href{trevor@bedford.io}{trevor@bedford.io}, \href{jbloom@fredhutch.org}{jbloom@fredhutch.org}}
}

\date{}

@@

\maketitle
\onehalfspacing

\begin{abstract}
Abstract text.
\end{abstract}

\section*{INTRODUCTION}
Very rough outline:
\begin{itemize}
\item Mutations are rampant in the evolution of human influenza virus. Seasonal H3N2 influenza virus in particular rapidly accumulates mutations.
\item The evolution of H3N2 is also characterized by clade competition and population turnover.
\item There have been efforts to predict evolutionary success.
\item Mutations that contribute to antigenic evolution largely determine strain success.
\item RNA viruses can accumulate deleterious mutations, and because the influenza does not appreciably recombine, deleterious mutations are linked to beneficial ones.
\item Mutations in HA that impact viral growth may influence strain success.
\item We need to understand the functional impact of mutations in HA.
\item Previously, we measured the effect of all possible single amino-acid mutations to an H1 hemagglutinin from the A/WSN/1933 (H1N1) strain~\citep{thyagarajan2014inherent,doud2016accurate}. 
\item However, this is a highly lab-adapted strain, and the measurements in this strain may not be particularly relevant for studying mutational processes of more contemporaneous strains circulating in the human population.
\item We chose to study the Perth/2009 H3 HA.
\item This also enabled a comparison of how the preferences have shifted for two diverged HA's.
\end{itemize}

\section*{RESULTS}
\label{sec:results}

\subsection*{Deep mutational scanning of HA from a recent strain of human H3N2 influenza virus}
We performed a deep mutational scan to measure the effects of all amino-acid mutations to the A/Perth/16/2009 (H3N2) HA on viral replication in cell culture. 
This strain was the H3N2 component of the influenza vaccine from 2010-2012.
We used a variant of this HA that supported better viral replication in cell culture than did the sequence in GenBank (Figure~\ref{suppfig:Perth2009_mut}).
This HA variant contained two mutations: G78D and T212I.
The G78D mutation occurs at low frequency in natural H3N2 sequences, and T212 is a site where a mutation to Ala rose to fixation in human influenza in $\sim$2011.
%We first generated virus carrying the wildtype Perth/2009 HA and NA and the WSN internal genes by reverse genetics, and serially passaged the virus in cell culture a total of six times.
%This passaging was completed to allow the virus to adapt to growth in cell culture and to find cell-culture adaptation mutations that emerged during the course of passaging.

We mutagenized the entire gene at the codon level to create mutant plasmid libraries harboring an average of $\sim$1.4 codon mutations per clone.
We then generated mutant virus libraries from the mutant plasmids using a helper virus system that enables the efficient generation of complex virus libraries~\citep{doud2016accurate} (Figure~\ref{fig:dms_overview}A).
In order to reduce biosafety concerns and to maximize viral titers, we used an HA-deficient helper virus carrying WSN/1933 internal and NA genes to rescue the mutant viruses.
Additionally, we used MDCK-SIAT1 cells constitutively expressing the TMPRSS2 protease, which facilitates HA cleavage and activation in the human airway~\citep{bottcher2006proteolytic, bottcher2010cleavage}.
All of the experiments were completed in full biological triplicate (Figure~\ref{fig:dms_overview}B). 
We also passaged and deep sequenced library 3 in technical replicate (denoted as library 3-1 and 3-2) to gauge to the amount of experimental noise occurring \textit{within} a single biological replicate.
As a control, we used the unmutated gene to generate and passage viruses carrying wildtype HA.

Deep sequencing revealed selection against non-functional HA variants as evidenced by the reduced mutation frequencies of the mutant viruses compared to their starting frequencies in the mutant plasmids (Figure~\ref{fig:dms_overview}C).
Specifically, stop codons were purged to 20-45\% of their original starting frequencies, after correcting for error rates from the wildtype controls.
Although the majority of stop codons were purged, incomplete purging of stop codons is likely due to complementation of virions. 
We also observed purging of nonsynonymous mutations to 30-40\% of their initial starting frequencies after error correction, suggesting strong selection against deleterious HA variants.

We next quantified the reproducibility of our deep mutational scanning measurements across biological and technical replicates. 
We inferred the amino-acid preferences at each site in the Perth/2009 HA using the method described in~\cite{bloom2015software} and implemented in the \texttt{dms\_tools2} software [\url{https://jbloomlab.github.io/dms_tools2/}].
These preferences represent an estimate of the 567 sites $\times$ 20 amino acids $=$ 11340 experimental measurements and are normalized to sum to one at each site.
The correlations of the amino-acid preferences between each pair of replicates is shown in Figure~\ref{fig:dms_overview}D.
The biological replicates are fairly well-correlated, with a Pearson's $R$ ranging from 0.69 to 0.78. 
Replicate 1 exhibits the least amount of correlation with the other biological replicates, consistent with the observation that this replicate showed the weakest selection against stop and nonsynonymous mutations and might therefore be subject to more experimental noise.
Of note, the two technical replicates 3-1 and 3-2 were only slightly more reproducible than that between biological replicates.
This suggests that bottlenecking of the virus library during the low MOI passage contributes to much of the noise observed in our experiments, as we are only able to passage a finite number of viral particles.

\begin{figure}
\centerline{\includegraphics[width=\textwidth]{figs/dms_overview/dms_overview.pdf}}
\caption{\label{fig:dms_overview}
{\bf Deep mutational scanning experiments of H3 HA.}
(A) We generated Perth/2009 HA mutant virus libraries using a helper virus approach described in~\cite{doud2016accurate}.
We then passaged this pool of mutant viruses at low MOI to establish a genotype-phenotype linkage and to select for functional HA variants. 
Deep sequencing of the variants before and after selection allowed us to infer the H3 site-specific amino-acid preferences.
(B) The DMS experiments were performed in biological triplicate, starting from independent preps of the wildtype HA genes to create the mutant plasmids. 
We also passaged and deep sequenced library 3 in technical replicate, denoted 3-1 and 3-2, to estimate the experimental noise within a biological replicate.
(C) Mutation frequencies of nonsynonymous, stop, and synonymous mutations for the mutant DNA, mutant virus, and unmutated DNA and virus controls. 
There is selection against nonsynonymous and stop codons in the mutant viruses. 
(D) The Pearson correlations for the amino-acid preferences between each pair of replicates are shown. 
The correlation between the technical replicates is comparable to that between biological replicates, indicating that much of the experimental noise comes from passaging at low MOI. 
}
\end{figure}

\subsection*{The HA preferences improve phylogenetic fit of natural H3 sequences}
How well do the Perth/2009 HA preferences inferred from experimental measurements describe the evolution of H3N2 influenza virus in nature?
This question can be addressed by evaluating how well experimentally informed codon substitution models (ExpCM's) constructed from our laboratory measurements improve phylogenetic fit of H3N2 evolution~\citep{hilton2017phydms}.
The results in Table~\ref{tab:phydms} show that the ExpCM outperforms conventional substitution models in describing the evolution of human H3N2 HA. 
The ExpCM also optimizes a stringency parameter for the preferences to more closely reflect the strength of selection in nature.
The stringency parameter in the ExpCM is equal to 2.44, which indicates that although the same amino acids are preferred, the strength of selection is more stringent in nature than in our experiments.
Figure~\ref{fig:logoplot} shows a logo plot of the Perth/2009 HA amino-acid preferences rescaled by this stringency parameter.

\begin{table}
\centering
\begin{tabular}{cccccccc}
\hline
\bf{Model} & \bf{$\Delta$AIC} & \bf{Log Likelihood} & \bf{Stringency} & \bf{$\omega$} & \bf{$\overline{\omega}$} & \bf{$\omega_{\alpha}$} & \bf{$\omega_{\beta}$} \\ \hline
ExpCM & 0.0 & -8439.33 & $2.44$ & $0.91$ & -- & -- & -- \\
Goldman-Yang M5 & 2166.06 & -9516.36 & -- & -- & $0.36$ & $0.30$ & $0.84$ \\
ExpCM, averaged across sites & 2504.18 & -9691.42 & $0.68$ & $0.32$ & -- & -- & -- \\
Goldman-Yang M0 & 2607.92 & -9738.29 & -- & $0.31$ & -- & -- & -- \\
\hline
\end{tabular}
\caption{\label{tab:phydms}
{\bf The site-specific amino-acid preferences are informative for describing human H3N2 evolution in nature.}
We implemented several codon substitution models for phylogenetic fitting of an alignment of human H3N2 HA sequences. 
The maximum likelihood values for each model were compared using the Akaike information criteria ($\Delta$AIC)~\citep{posada2004model}.
An experimentally-informed codon substitution model (ExpCM) built from the preferences averaged across all replicates performs better than conventional substitution models, specifically the M0 and M5 models in~\cite{yang2000codon}.
A non-site-specific ExpCM informed by preferences averaged across all sites performs comparably to the GY94 class of models, indicating the importance of site-specificity in the ExpCM.
The optimized parameters for each model are also shown.
}
\end{table}

A closer examination of the logo plot reveals that the preferences generally agree with existing knowledge about HA's biochemistry.
For instance, sites that form structurally important disulfide bridges (sites 52 \& 277, 64 \& 76, 97 \& 139, 281 \& 305, 14 \& 137-HA2, 144-HA2 \& 148-HA2)~\citep{waterfield1981disulphide} possess high preference for cysteine.
At residues involved in receptor binding, there are strong preferences for the amino acids at sites Y98, D190, W153, and S228~\citep{weis1988structure,martin1998studies}.
A positively charged amino acid at site 329 is important for cleavage activation of the HA0 precursor, and indeed this site exhibits a high preference for arginine~\citep{kido1992isolation, stech2005new}.
However, a notable exception occurs at the start codon at position -16, which does not show a strong preference for methionine. 
This codon is part of the signal peptide and is cleaved from the mature HA protein.
One possible explanation for why we do not see a strong preference for Met at this site is due to alternative translation initiation occurring at a downstream or upstream start site, as has been described for HA~\citep{girard2011upstream}.
% No upstream or downstream "ATG"'s are present. I did find an in-frame "ATC" one codon upstream of the canonical start site (in the non-coding region), and interestingly, all Perth/2009 HA sequences deposited in GenBank have an "ATT" in this non-coding position, including Seema's Perth HA.
% I6M(HA2) mutant slightly increases the fusion pH (Daniels 1985 Cell)
% We see high preference for T,S,Q over the wildtype Ile at site 226. Although Q226 in an H3 background is characterized as having preference for avian-type receptors, most of these studies were done in older H3 strains such as Aichi. Wu 2017 Cell Host Microbe has shown extensive epistasis in the RBS, specifically in the 220-loop, so perhaps the amino acids tolerated in this loop have shifted over time. Indeed the wildtype amino acids at 225 and 226 are different in Perth compared to HK68 and Aichi.

\begin{figure}
\centerline{\includegraphics[width=\textwidth]{figs/prefslogoplot/rescaled-avgprefs_prefs.pdf}}
\caption{\label{fig:logoplot}
{\bf The site-specific amino-acid preferences of the Perth/2009 HA.}
This logoplot shows the site-specific amino-acid preferences for the averaged replicates rescaled by the stringency parameter (Table 1) estimated by phydms.
The height of each letter is proportional to its preference at that site, and the preferences for all sites are normalized to sum to 1.
The sites are in H3 numbering.
The top overlay bar shows the relative solvent accessibility.
The bottom overlay bar is colored by the HA domain (sig. pep. = signal peptide, HA1 ecto. = HA1 ectodomain, HA2 ecto. = HA2 ectodomain, TM = transmembrane domain, cyto. tail. = cytoplasmic tail).
The letters directly above each logo indicate the wildtype amino acid at that site.
}
\end{figure}

\subsection*{The mutational tolerance of the head and the stalk shows less of a contrast in H3 than in H1}
We next sought to investigate the inherent mutational tolerance of the Perth/2009 HA. 
Figure~\ref{fig:mut_tolerance} shows the mutational tolerance as calculated from the rescaled Perth/2009 H3 preferences and the rescaled WSN/1933 H1 preferences mapped onto the HA crystal structures.
We found antigenic site C and the most distal portion of the globular head near the 190-helix in the Perth/2009 H3 to be tolerant of mutations.
Interestingly, the H3 stalk including the shorter $\alpha$-helix (helix A) is relatively mutationally tolerant compared to the tolerance of the globular head domain. 
This observation suggests that the stalk may be prone to escape from antibodies, and agrees with previous work demonstrating that it is possible to select for antigenic mutants in H3 by broadly-neutralizing stalk-targeting antibodies~\citep{ekiert2011highly, friesen2014common, chai2016two, yamayoshi2017human}.
% In contrast, the head domain of the WSN/1933 H1 is more mutationally tolerant relative to its stalk domain.

The sites inside the receptor binding pockets are highly functionally constrained and were found to be relatively mutationally intolerant in both H3 and H1~\citep{wilson1981structure}.
In contrast, the residues surrounding the receptor binding pocket are fairly mutationally tolerant, which may contribute to antigenic evolution as these sites are under strong immune pressure.~\citep{wiley1981structural}.

\begin{figure}
\centerline{\includegraphics[width=\textwidth]{figs/mut_tolerance/entropy_heatmap.pdf}}
\caption{\label{fig:mut_tolerance}
{\bf Mutational tolerance mapped onto each site of HA.}
Mutational tolerance as calculated by the Shannon entropy of a given site's amino-acid preferences are mapped onto the structure of the H3 trimer (PDB 4O5N;~\cite{lee2014receptor}) and the H1 trimer (PDB 1RVX;~\cite{gamblin2004structure}), with both trimers in approximately the same orientation. 
The site entropies were calculated from the preferences measured in the Perth/2009 H3 (left panel) from this study, or the preferences measured in the WSN/1933 H1 (right panel) from~\cite{doud2016accurate}. 
Lighter shades of blue or red signify low mutational tolerance, while darker shades of blue or red signify high mutational tolerance. 
For each HA, the structure on the left side colors the full HA trimer, while the structure on the right side colors only one of the monomers.
The sialic acid receptor is shown as black sticks.
The Perth/2009 H3 shows relatively high mutational tolerance in the stalk region, particularly in helix A, compared to the head region. 
High mutational tolerance in H3 was also observed near antigenic site C and near the 190 helix.
The head region of the WSN/1933 H1 is mutationally tolerant compared to the relatively intolerant stalk region. 
}
\end{figure}

\subsection*{The experimental measurements can help discriminate successful influenza virus lineages}
A challenge in vaccine strain selection is predicting which strain will dominate the upcoming influenza season.
An important question to thereby address is if our experimental measurements are useful in differentiating between strains of human H3N2 influenza virus that have succeeded and those that have died out.
To investigate this, we used our preference dataset and an H3N2 phylogeny from 1968-2012 (Figure~\ref{fig:trunkvssidebranch}A) to calculate the effects of mutations for the evolutionarily successful trunk lineage and for side branches which have died out.
We found that strains with mutations measured to be more beneficial to viral growth in our experiments tend to succeed in nature.
Figure~\ref{fig:trunkvssidebranch}B shows the effects of trunk and side branch mutations in five-year intervals for every year from 1968-2012. 
On average, trunk mutations are towards more preferred amino acids compared to side branch mutations, and this was true for all intervals.
Importantly, trunk mutations are significantly more favorable than side branch mutations when calculating the effects from the entirety of the phylogenetic tree (Figure~\ref{fig:trunkvssidebranch}C).

Because tip nodes can contain egg- or cell-passaged isolates~\citep{wu2017structural,mcwhite2016sequence,skowronski2016mutations} and our experiments were performed in cell culture, we examined the effects of mutations on terminal side branch nodes to see if these would rank more highly than internal node mutations.
Instead, terminal node mutations are on average towards less preferred amino acids than internal node mutations (Figure~\ref{fig:trunkvssidebranch}C), and both internal and terminal node mutations are significantly less favorable than those on the trunk.
Therefore, strains that have accumulated mutations that we experimentally measured to be unpreferred generally die out in nature, while more favored mutations provide a selective advantage to the trunk.
These findings demonstrate the importance of the functional impacts of HA mutations in determining strain success.

\begin{figure}
\centerline{\includegraphics[width=0.75\textwidth]{figs/trunkvssidebranch/trunkvssidebranch.pdf}}
\caption{\label{fig:trunkvssidebranch}
{\bf Mutations in evolutionary successful strains tend to be more favorable than in strains that die out.}
(A) Phylogenetic tree of human H3N2 influenza virus from 1968-present. 
The trunk is shown in red, and side branches are shown in blue.
The gray branches represent the part of the tree for which we cannot yet distinguish the trunk from side branches.
(B) Using the Perth/2009 H3 preferences, we calculated the log$_{2}$ mutational effect for trunk and side branch mutations in windows of 5 years for every year from 1968-2013. 
The median log$_{2}$ mutational effect in a given window is shown as circles for trunk mutations and triangles for side branch mutations. 
The shaded region demarcates the interquartile range of trunk and side branch mutational effects.
Negative numbers signify mutations towards less preferred amino acids, while positive numbers signify more preferred mutations.
The median trunk mutational effects are consistently higher than the median side branch mutational effects for all windows.
(C) The log$_{2}$ mutational effect for all side branch and all trunk mutations (left panel), in addition to all mutations in internal nodes and terminal nodes on the side branches (right panel) are shown.
The preferences were randomized 10,000 times to estimate significance.
The effects of trunk mutations are higher than side branch internal and terminal node mutations.
}
\end{figure}

How distantly can the preferences be extended to describe differences in successful and unsuccessful strains?
To explore this question, we scored the complete HA sequence of every node in the tree using the preferences by quantifying a \textit{sequence preference} metric.
The sequence preference is defined as 
$$\displaystyle\sum_{r} \ln \pi_{r, a}$$
where $\pi_{r, a}$ is the preference for amino acid $a$ at site $r$.
Consistent with the finding that trunk mutations are generally more favorable than side branch mutations, Figure~\ref{fig:sequence_preference} shows that the sequences of trunk nodes also tend to be more highly preferred than those of side branch nodes.
Interestingly, the sequence preferences increase over time as the nodes approach the Perth/2009 strain and its closely related nodes, which all exhibit sequence preferences that are higher than that of the trunk. 
The observation that the strain in which we performed our deep mutational scan has one of the highest sequence preferences illustrates epistatic interactions among mutations such that an unpreferred mutation in one background may be preferred in another.
% Due to the sheer number of possible pairwise combinations of mutations, it is impossible for us to experimentally characterize these epistatic interactions among mutations.

\comment{this section will be fleshed out more once I figure out how to plot the substitutions from the root at epitope / non-epitope sites, as per Trevor's suggestion.}
We reasoned that much of the increase in sequence preference toward Perth/2009 could be attributed to epitope sites, as most substitutions have occurred at sites under strong immune pressure.
We therefore assessed each node's preference at epitope and non-epitope sites as defined by~\cite{wolf2006long}.
Indeed, the preferences at epitope sites resembles those for all HA sites (Figure~\ref{fig:sequence_preference}), with preferences increasing over time.
At non-epitope sites, however, the trunk preferences remain relatively constant while side branch preferences tend to drop below the trunk.
These observations highlight the extensive epistasis among epitope sites, as has even been noted for sites within the 220-loop surrounding the receptor binding pocket~\citep{wu2017diversity}.
Many successful epitope mutations are likely contingent upon the background in which they arise.
Yet, we are able to distinguish trunk sequences as more favorable than side branch sequences, indicating that the preferences are still of utility over short evolutionary timescales.

\begin{figure}
\centerline{\includegraphics[width=0.75\textwidth]{figs/sequence_preference/sequence_preference.pdf}}
\caption{\label{fig:sequence_preference}
{\bf The HA sequences of trunk nodes tend to be more preferred than those of side branch nodes.}
We used the preferences to calculate the HA sequence preference normalized by the number of sites of every node in a human H3N2 phylogenetic tree.
To calculate the sequence preference, we used the entire HA sequence (top panel), only epitope sites (center panel) or non-epitope sites (bottom panel) as defined by~\cite{wolf2006long}.
Higher preferences are closer to zero while lower preferences are more negative.
The preferences at all sites and at epitope sites increase as the tree approaches the Perth/2009 strain, indicating epistasis among epitope sites.
The trunk nodes generally exhibit higher sequence preferences than do side branch nodes.
}
\end{figure}

Can we then distinguish lineage-specific mutational effects using the preferences measured in a distantly related HA homolog?
We used the preferences measured previously in the WSN/1933 H1~\citep{doud2016accurate} to quantify the effects of H3 trunk and side branch mutations, shown in Figure~\ref{fig:WSN_trunkvssidebranch}.
It is evident that we do not see trunk mutations significantly more favored than side branch mutations, suggesting that our ability to discriminate successful and unsuccessful strains degrades over sufficiently long evolutionary distances.

\begin{figure}
\centerline{\includegraphics[width=\textwidth]{figs/WSN_trunkvssidebranch/WSN_trunkvssidebranch.pdf}}
\caption{\label{fig:WSN_trunkvssidebranch}
{\bf The WSN/1933 H1 preferences do not reveal differences in trunk vs side branch mutational effects}
(A) We used the WSN/1933 H1 preferences from~\cite{doud2016accurate} to calculate the log$_{2}$ mutational effects of trunk and side branch mutations from the inferred H3N2 phylogeny in Figure~\ref{fig:trunkvssidebranch}.
There is not a significant difference in trunk vs side branch mutational effects.
(B) We also performed a sliding window analysis with the WSN/1933 H1 preferences.
There is not a distinct difference in trunk and side branch mutational effects.
}
\end{figure}

\subsection*{The H1 and H3 preferences at many sites have shifted}
How shifted are the preferences between evolutionarily distant homologs such as H1 and H3?
Although we have previously compared the preferences between related protein homologs of NP~\citep{doud2015site} and of Env \comment{cite haddox2017}, the HA homologs we have experimentally studied are considerably more diverged than either of these pairs.
The WSN/1933 H1 and the Perth/2009 H3 fall into two phylogenetically distinct groups and share 42\% amino-acid identity (Figure~\ref{fig:distance_distribution}A), providing an opportunity to analyze what has shifted and what has remained conserved.
Simply correlating the preferences between H1 and H3 reveals that the replicate measurements within a homolog are more correlated than between homologs (Figure~\ref{fig:distance_distribution}B).

>>>>

\documentclass[9pt,twocolumn,twoside]{pnas-new}
% Use the lineno option to display guide line numbers if required.
% Note that the use of elements such as single-column equations
% may affect the guide line number alignment.

\templatetype{pnasresearcharticle} % Choose template 
% {pnasresearcharticle} = Template for a two-column research article
% {pnasmathematics} = Template for a one-column mathematics article
% {pnasinvited} = Template for a PNAS invited submission

@@

% Use letters for affiliations, numbers to show equal authorship (if applicable) and to indicate the corresponding author
\author[a,d,e,1]{Juhye M. Lee}
\author[b,f,1]{John Huddleston} 
\author[a,d,e]{Michael B. Doud}
\author[a,f]{Kathryn A. Hooper}
\author[b,c,2]{Trevor Bedford}
\author[a,c,d,2]{Jesse D. Bloom}

\affil[a]{Basic Sciences Division}
\affil[b]{Vaccine and Infectious Diseases Division}
\affil[c]{and Computational Biology Program, Fred Hutchinson Cancer Research Center, Seattle, WA 98109, USA}
\affil[d]{Department of Genome Sciences}
\affil[e]{Medical Scientist Training Program}
\affil[f]{and Molecular and Cellular Biology Program, University of Washington, Seattle, WA 98195, USA}

% Please give the surname of the lead author for the running footer
\leadauthor{Lead author last name}

% Please add here a significance statement to explain the relevance of your work
\significancestatement{Authors must submit a 120-word maximum statement about the significance of their research paper written at a level understandable to an undergraduate educated scientist outside their field of speciality. The primary goal of the Significance Statement is to explain the relevance of the work in broad context to a broad readership. The Significance Statement appears in the paper itself and is required for all research papers.}

% Please include corresponding author, author contribution and author declaration information
\authorcontributions{Please provide details of author contributions here.}
\authordeclaration{Please declare any conflict of interest here.}
\equalauthors{\textsuperscript{1}J.M.L. and J.H. contributed equally to this work.}
\correspondingauthor{\textsuperscript{2}To whom correspondence should be addressed. E-mail: trevor@bedford.io, jbloom@fredhutch.org}

% Keywords are not mandatory, but authors are strongly encouraged to provide them. If provided, please include two to five keywords, separated by the pipe symbol, e.g:
\keywords{Keyword 1 $|$ Keyword 2 $|$ Keyword 3 $|$ ...}

\begin{abstract}
Please provide an abstract of no more than 250 words in a single paragraph. Abstracts should explain to the general reader the major contributions of the article. References in the abstract must be cited in full within the abstract itself and cited in the text.
\end{abstract}

\dates{This manuscript was compiled on \today}
\doi{\url{www.pnas.org/cgi/doi/10.1073/pnas.XXXXXXXXXX}}

@@

% Optional adjustment to line up main text (after abstract) of first page with line numbers, when using both lineno and twocolumn options.
% You should only change this length when you've finalised the article contents.
\verticaladjustment{-2pt}

\maketitle
\thispagestyle{firststyle}
\ifthenelse{\boolean{shortarticle}}{\ifthenelse{\boolean{singlecolumn}}{\abscontentformatted}{\abscontent}}{}

% If your first paragraph (i.e. with the \dropcap) contains a list environment (quote, quotation, theorem, definition, enumerate, itemize...), the line after the list may have some extra indentation. If this is the case, add \parshape=0 to the end of the list environment.
\dropcap{T}his PNAS journal template is provided to help you write your work in the correct journal format.  Instructions for use are provided below.

Note: please start your introduction without including the word ``Introduction'' as a section heading (except for math articles in the Physical Sciences section); this heading is implied in the first paragraphs.

\section*{Results}
\label{sec:results}

\subsection*{Deep mutational scanning of HA from a recent strain of human H3N2 influenza virus}

\subsection*{Submitting Manuscripts}

All authors must submit their articles at \href{http://www.pnascentral.org/cgi-bin/main.plex}{PNAScentral}. If you are using Overleaf to write your article, you can use the ``Submit to PNAS'' option in the top bar of the editor window.

\subsection*{Format}

Many authors find it useful to organize their manuscripts with the following order of sections;  Title, Author Affiliation, Keywords, Abstract, Significance Statement, Results, Discussion, Materials and methods, Acknowledgments, and References. Other orders and headings are permitted.

\subsection*{Manuscript Length}

PNAS generally uses a two-column format averaging 67 characters, including spaces, per line. The maximum length of a Direct Submission research article is six pages and a PNAS PLUS research article is ten pages including all text, spaces, and the number of characters displaced by figures, tables, and equations.  When submitting tables, figures, and/or equations in addition to text, keep the text for your manuscript under 39,000 characters (including spaces) for Direct Submissions and 72,000 characters (including spaces) for PNAS PLUS.

\subsection*{References}

References should be cited in numerical order as they appear in text; this will be done automatically via bibtex, e.g. \cite{belkin2002using} and \cite{berard1994embedding,coifman2005geometric}. All references should be included in the main manuscript file.

\subsection*{Data Archival}

PNAS must be able to archive the data essential to a published article. Where such archiving is not possible, deposition of data in public databases, such as GenBank, ArrayExpress, Protein Data Bank, Unidata, and others outlined in the Information for Authors, is acceptable.

\subsection*{Language-Editing Services}
Prior to submission, authors who believe their manuscripts would benefit from professional editing are encouraged to use a language-editing service (see list at www.pnas.org/site/authors/language-editing.xhtml). PNAS does not take responsibility for or endorse these services, and their use has no bearing on acceptance of a manuscript for publication.

%\begin{figure}%[tbhp]
%\centering
%\includegraphics[width=.8\linewidth]{frog}
%\caption{Placeholder image of a frog with a long example caption to show justification setting.}
%\label{fig:frog}
%\end{figure}

%\begin{SCfigure*}[\sidecaptionrelwidth][t]
%\centering
%\includegraphics[width=11.4cm,height=11.4cm]{frog}
%\caption{This caption would be placed at the side of the figure, rather than below it.}\label{fig:side}
%\end{SCfigure*}

\subsection*{Digital Figures}
\label{sec:figures}

Only TIFF, EPS, and high-resolution PDF for Mac or PC are allowed for figures that will appear in the main text, and images must be final size. Authors may submit U3D or PRC files for 3D images; these must be accompanied by 2D representations in TIFF, EPS, or high-resolution PDF format.  Color images must be in RGB (red, green, blue) mode. Include the font files for any text.

Figures and Tables should be labelled and referenced in the standard way using the \verb|\label{}| and \verb|\ref{}| commands.

Figure \ref{fig:frog} shows an example of how to insert a column-wide figure. To insert a figure wider than one column, please use the \verb|\begin{figure*}...\end{figure*}| environment. Figures wider than one column should be sized to 11.4 cm or 17.8 cm wide. Use \verb|\begin{SCfigure*}...\end{SCfigure*}| for a wide figure with side captions.

\subsection*{Single column equations}

Authors may use 1- or 2-column equations in their article, according to their preference.

To allow an equation to span both columns, options are to use the \verb|\begin{figure*}...\end{figure*}| environment mentioned above for figures, or to use the \verb|\begin{widetext}...\end{widetext}| environment as shown in equation \ref{eqn:example} below.

Please note that this option may run into problems with floats and footnotes, as mentioned in the \href{http://texdoc.net/pkg/cuted}{cuted package documentation}. In the case of problems with footnotes, it may be possible to correct the situation using commands \verb|\footnotemark| and \verb|\footnotetext|.

%% Do not use widetext if paper is in single column.
\begin{widetext}
\begin{align*}
(x+y)^3&=(x+y)(x+y)^2\\
       &=(x+y)(x^2+2xy+y^2) \numberthis \label{eqn:example} \\
       &=x^3+3x^2y+3xy^3+x^3. 
\end{align*}
\end{widetext}

\begin{table}%[tbhp]
\centering
\caption{Comparison of the fitted potential energy surfaces and ab initio benchmark electronic energy calculations}
\begin{tabular}{lrrr}
Species & CBS & CV & G3 \\
\midrule
1. Acetaldehyde & 0.0 & 0.0 & 0.0 \\
2. Vinyl alcohol & 9.1 & 9.6 & 13.5 \\
3. Hydroxyethylidene & 50.8 & 51.2 & 54.0\\
\bottomrule
\end{tabular}

\addtabletext{nomenclature for the TSs refers to the numbered species in the table.}
\end{table}

\subsection*{Supporting Information (SI)}

The main text of the paper must stand on its own without the SI. Refer to SI in the manuscript at an appropriate point in the text. Number supporting figures and tables starting with S1, S2, etc. Authors are limited to no more than 10 SI files, not including movie files. Authors who place detailed materials and methods in SI must provide sufficient detail in the main text methods to enable a reader to follow the logic of the procedures and results and also must reference the online methods. If a paper is fundamentally a study of a new method or technique, then the methods must be described completely in the main text. Because PNAS edits SI and composes it into a single PDF, authors must provide the following file formats only.

\subsubsection*{SI Text}

Supply Word, RTF, or LaTeX files (LaTeX files must be accompanied by a PDF with the same file name for visual reference).

\subsubsection*{SI Figures}

Provide a brief legend for each supporting figure after the supporting text. Provide figure images in TIFF, EPS, high-resolution PDF, JPEG, or GIF format; figures may not be embedded in manuscript text. When saving TIFF files, use only LZW compression; do not use JPEG compression. Do not save figure numbers, legends, or author names as part of the image. Composite figures must be pre-assembled.

\subsubsection*{3D Figures}

Supply a composable U3D or PRC file so that it may be edited and composed. Authors may submit a PDF file but please note it will be published in raw format and will not be edited or composed.

\subsubsection*{SI Tables}

Supply Word, RTF, or LaTeX files (LaTeX files must be accompanied by a PDF with the same file name for visual reference); include only one table per file. Do not use tabs or spaces to separate columns in Word tables.

\subsubsection*{SI Datasets}

Supply Excel (.xls), RTF, or PDF files. This file type will be published in raw format and will not be edited or composed.

\subsubsection*{SI Movies}

Supply Audio Video Interleave (avi), Quicktime (mov), Windows Media (wmv), animated GIF (gif), or MPEG files and submit a brief legend for each movie in a Word or RTF file. All movies should be submitted at the desired reproduction size and length. Movies should be no more than 10 MB in size.

\subsubsection*{Still images}

Authors must provide a still image from each video file. Supply TIFF, EPS, high-resolution PDF, JPEG, or GIF files.

\subsubsection*{Appendices}

PNAS prefers that authors submit individual source files to ensure readability. If this is not possible, supply a single PDF file that contains all of the SI associated with the paper. This file type will be published in raw format and will not be edited or composed.

\matmethods{Please describe your materials and methods here. This can be more than one paragraph, and may contain subsections and equations as required. Authors should include a statement in the methods section describing how readers will be able to access the data in the paper.

\subsection*{Subsection for Method}
Example text for subsection.
}

\showmatmethods{} % Display the Materials and Methods section

\acknow{Please include your acknowledgments here, set in a single paragraph. Please do not include any acknowledgments in the Supporting Information, or anywhere else in the manuscript.}

\showacknow{} % Display the acknowledgments section

% \pnasbreak splits and balances the columns before the references.
% Uncomment \pnasbreak to view the references in the PNAS-style
% If you see unexpected formatting errors, try commenting out \pnasbreak
% as it can run into problems with floats and footnotes on the final page.
%\pnasbreak

% Bibliography
\bibliography{pnas-sample}
